---
title: 'Practical Machine Learning: predicting quality activity'
author: "LOGp"
date: "17/08/2014"
output: html_document
---

```{r}
library(ggplot2)
library(caret)
```

## Executive summary



## Download and load data
### Download
```{r cache=TRUE}
train_url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
train_file = "train.csv"
test_url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
test_file = "test.csv"

#download.file(train_url, train_file, "curl")
#download.file(test_url, test_file, "curl")
```
### Load data into R
```{r cache=TRUE}
train = read.csv(train_file)
test = read.csv(test_file)
```

### Plot classe data
Just to see if the classes are balanced:
```{r}
qplot(classe, data=train)
```

### Clean data
Check the data types of data, and cast numerical values to numeric class.
```{r}
for (col in c("kurtosis_roll_belt", "kurtosis_picth_belt", "kurtosis_yaw_belt", "skewness_roll_belt",
              "skewness_roll_belt.1", "skewness_yaw_belt", "max_yaw_belt", "min_yaw_belt",
              "amplitude_yaw_belt", "kurtosis_roll_arm", "kurtosis_picth_arm", "skewness_yaw_arm",
              "kurtosis_roll_dumbbell", "kurtosis_picth_dumbbell", "kurtosis_yaw_dumbbell", 
              "skewness_roll_dumbbell", "skewness_pitch_dumbbell", "skewness_yaw_dumbbell", 
              "max_yaw_dumbbell", "min_yaw_dumbbell", "amplitude_yaw_dumbbell", "kurtosis_roll_forearm", 
              "kurtosis_picth_forearm", "kurtosis_yaw_forearm", "skewness_roll_forearm", 
              "skewness_pitch_forearm", "skewness_yaw_forearm", "max_yaw_forearm", 
              "min_yaw_forearm", "amplitude_yaw_forearm", "kurtosis_yaw_arm", "skewness_roll_arm",
              "skewness_pitch_arm")) {
  train[[col]] = as.numeric(as.character(train[[col]]))
}
```

### Define functions for prediction

Remove features with unique values
```{r}
remove_unary_features = function (X) {
  cardinality = sapply(X, function(x) {length(unique(x))})
  X[, ! cardinality == 1]
}

```

Populate NA values:
* median for numeric and values
```{r}
set_NA = function (x) {
  if (class(x) %in% c("numeric", "integer")) {
    m = median(x, na.rm=TRUE)
    if (class(x) == "integer") {
      m = as.integer(m)
    }
    x[is.na(x)] = m
    x
  }
  else {
    x
  }
}

populate_NAs = function (X) {
  as.data.frame(lapply(X, set_NA))
}
```



## Prediction model
I will train a mutliple logistic regression model, and evaluate the prediction error via cross-validation.
For that purpose, I will perform 5 regressions, for each class value, and then select the classe as the  one with the hightest probability.



### Cross validation
Define 5-folds cross-validation groups:
```{r}
k=5
set.seed(1324)
folds = createFolds(train$classe, k=k, list=TRUE)
```


Run cross-validation
```{r}
#for (i in 1:k) {
i=k
  data.train = train[-folds[[i]], ]
  # These 2 function calls must be decomposed in preproces to apply them to test data
  data.train = remove_unary_features(data.train)
  data.train = populate_NAs(data.train)
  tree.fit = train(classe ~ ., model="rpart", data=data.train)
  tree.predicted = predict(tree.fit, newdata=data.test)

  lm.fit = lm(classe ~ ., data=data.train)

  data.test = train[folds[[i]], ]
  # To be removed
  data.test = remove_unary_features(data.test)
#}
```










